---
title: "Practical Machine Learning - Prediction Assignment Writeup"
author: "gjamnitz"
date: "Sunday, May 24, 2015"
output: html_document
---

##Introduction
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project the goal was to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. 

##Loading data
Empty and '#DIV/0!' values were translated to NA's.
```{r}
suppressMessages(suppressWarnings(library(caret)))
suppressMessages(suppressWarnings(library(randomForest)))
training <- read.csv("pml-training.csv", na.strings=c("NA","#DIV/0!",""))
testing <- read.csv("pml-testing.csv", na.strings=c("NA","#DIV/0!",""))
```

##Preprocessing data
Removing unnecessary (misleading) variables and summary rows (where *new_window == true*).
```{r}
training <- training[training$new_window == 'no', ]
training$X <- NULL
training$user_name <- NULL
training$raw_timestamp_part_1 <- NULL
training$raw_timestamp_part_2 <- NULL
training$cvtd_timestamp <- NULL
training$num_window <- NULL
training$new_window <- NULL
```

Removing all NA variables from training set (only 53 columns remained)
```{r}
notEmptyCols <- colSums(is.na(training)) != nrow(training)
training <- training[, notEmptyCols]
dim(training)
```

##Cross validation 
Splitting the cleaned training set into 2 parts.
```{r}
inTrain <- createDataPartition(y = training$classe, p=0.6, list=FALSE)
trainingTrain <- training[inTrain, ]
trainingTest <- training[-inTrain, ]
dim(trainingTrain)
dim(trainingTest)
```

## Building the model
I chose Random Forest model for prediction.
```{r}
modelRf <- train(y = as.factor(trainingTrain$classe), x = trainingTrain[, -53], method = "rf",
                  tuneGrid=data.frame(mtry=3), trControl = trainControl(method="none"))

```
We can see that sensitivity and specificity are both high and confusion matrix is convincing as well.

## The final model
```{r}
modelRf$finalModel
```
The model tried 500 trees. The confusion matrix shows that this model is quite efficient.

The following graph shows the relative importance of the variables in final model:
```{r fig.width=9, fig.height=7}
plot(varImp(modelRf))
```


## Checking model accuracy with the training subset
```{r}
confusionMatrix(predict(modelRf, trainingTest), trainingTest$classe)
```
Accuracy is near to 100% and the confusion matrix seems to be almost excellent.

##Conclusion
Random Forest model was a good choice for predicting Weight Lifting Exercise types. 
